---
title: "Exercise1_TP1"
author: "Emily Schmidt"
date: '2022-10-23'
output: html_document
---

### Problem 3.4 - Laptop Sales at a London Computer Chain: Interactive Visualization  
The file LaptopSales.txt is a comma-separated file with nearly 300,000 rows. ENBIS (the European Network for Business
and Industrial Statistics) provided these data as part of a contest organized in the fall of 2009.  
**Scenario**: Imagine that you are a new analyst for a company called Acell (a company selling laptops). You have been provided with data about products and sales. You need to help the company with their business goal of planning a product strategy and pricing policies that will maximize Acell’s projected revenues in 2009. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
library(data.table) # Fast aggregation of large data sets and a fast file reader
library(lubridate) # Commute date-times
library(ggplot2) # Provides helpful commands to create complex plots from data in a data frame
library(reshape2) # Easily transform our data into whatever structure we may need
library(dplyr) # Data manipulation operations such as applying filter
library(tidyr) # Contains tools for changing the shape and hierarchy of a data set
library(caret) # Functions that attempt to streamline the process for creating predictive models
library(corrplot) # Graphical display of a correlation matrix, confidence interval or general matrix
library(lattice) # An implementation of Trellis graphics 
library(tidyverse) # Transform data
library(scales) # How to normalize data
library(ggmap) # Maps visualizations
library(osmdata) # Open street maps
library(sp) # Spatial data
library(rgdal) # Needed due to error with the map
library(UpSetR) # Creates visualizations of intersecting sets using a novel matrix design, along with visualizations of several common set
library(forcats) # Need to use fct_infreq in a bar chart to sort
library(cowplot) # Arrange plots into a grid and label them
```

### Preliminary data exploratory:
```{r}
setwd("C:\\Users\\emann\\GSEM Master\\Data Mining\\Homework 1") # Set directory
getwd() # Check the working directory

Laptops <- fread("LaptopSales_red.csv", sep = ",", header = T) # Load your data, Cereals.csv

dim(Laptops) # Get or set the dimension of the specified matrix, array or data frame
head(Laptops) # Display the first n rows present in the input data frame
View(Laptops) # Invoke a spreadsheet-style data viewer within RStudio
summary(Laptops) # Produce result summaries of the results of various model fitting functions
str(Laptops) # Compactly displaying the internal structure of a R object

head(which(is.na(Laptops))) # Function returns the positions with missing values
sum(is.na(Laptops))  # Sum the total amount of missing values in the data set
Laptops <- na.omit(Laptops) # Remove the missing values
summary(Laptops) # Call the data set to ensure that missing values are removed
```
One of the most important parts of analyzing your data is to conduct a preliminary exploratory review. The following points outline high-level conclusions:  
- There are 148,786 observations and 17 variables within the raw data set   
- There is information on date, configuration, customer/store location, and laptop features  
- When calling summary() and str(), we notice there are missing values and various types of variables (chr, int, and num)  
- Diving deeper into the missing values, there are 6,826 NAs. Therefore, we omit. When recalling summary(), there now should be a total of 142,049 observations that will we proceed with using throughout this analysis.  

**A** - Price Questions:   
**i.** *At what price are laptops actually selling?*
```{r}
summary(Laptops$Retail.Price) # Produce result summaries of the Retail.Price 

# Histogram shows the distribution of the outcome variable
rp.hist <- ggplot(Laptops, aes(x = Retail.Price)) +
  geom_histogram(color="white", fill="black",binwidth = 40, linetype = "dashed") + # Graph design
  ggtitle("Distirbution of Retail Price") + # Title name
  ylab("Count") + xlab("Retail Price")  + # Label names
  theme(plot.title = element_text(size=50)) + # Title text size
  theme_classic(base_size = 8) # Elongating the graph to show all values

rp.hist

#Box plot shows the distribution quartiles, min, max, median, etc.
rp.box <- ggplot(Laptops, aes(x="", y=Retail.Price)) + 
  geom_boxplot(outlier.shape = 1, fill = "NA") +
  stat_boxplot(geom ='errorbar', width = .5) + # Added the horizontal bars for visualization
  ggtitle("Distirbution of Retail Price") + # Title name
  ylab("Count") + xlab("Retail Price")  + # Label names
  stat_summary(fun="mean",color="red",geom = "errorbar", aes(ymax = ..y.., ymin = ..y..), linetype = "dashed", size = 1) + # Show difference between mean and median
  theme(plot.title = element_text(size=50)) + # Title text size
  theme_classic(base_size = 8) # Elongating the graph to show all values

rp.box

min(Laptops$Retail.Price) # Calculate minimum
max(Laptops$Retail.Price) # Calculate maximum
mean(Laptops$Retail.Price) # Calculate mean
median(Laptops$Retail.Price) # Calculate median
```

The retail prices range from under $200 to over $850, averaging around $508. Customers typically will purchase a laptop that is slightly under $500 though. Overall, both the histogram and the box plot assume to be normally distributed. Outside of the minimum and maximum values, one can see the numerous outliers that are present within the data.

**ii.** *Does price change with time?* 
```{r, message=FALSE, warning=FALSE}
Dates <- as.Date(Laptops$Date, "%m/%d/%Y") # Transform the date into M-D-Y
tab = data.frame(Dates,Laptops$Retail.Price) # Store new Date and Retail.Price

tab$Month = as.Date(cut(Dates,breaks = "month")) # Divide the Date into different ranges - months in this case
head(tab$Month) # Display the first n rows present in the input data frame

tab$Week = as.Date(cut(Dates,breaks = "week",start.on.monday = F)) # Date broken into weeks, starting on Monday
head(tab$Week)

tab$Days = as.Date(cut(Dates,breaks = "day",start.on.monday = F)) # Date broken into days, starting on Monday
head(tab$Days)

Num = table(Laptops$Retail.Price) # Creates a table with all Retail Prices
head(Num) # Check 'Num'
Laptops2 = cbind(Laptops,Num) # Mergers data together
Laptops2 <- na.omit(Laptops2) # Remove the missing values
summary(Laptops2) # Check 'Laptops2'

bymonth = aggregate(cbind(Retail.Price)~month(Dates),data = Laptops2, FUN = mean) # Creates table with Retail Price and month
bymonth # Calls the table

Price = Laptops2$Retail.Price # Rename

# Line chart based on monthly data
ggplot(data = tab,aes(Month,Price)) +
        stat_summary(fun = median,geom = "line") +
        scale_x_date(labels=date_format("%Y-%m"),breaks = "1 month") +
        ggtitle("Monthly Price Evolution") +
        theme(plot.title = element_text(hjust = 0.5)) + # Adjust plot title
        theme(axis.text.x = element_text(face = "bold",color = "black",size = 8)) # Adjust x axis title

# Line chart based on weekly data
ggplot(data = tab,aes(Week,Price),na.rm = T) +
        stat_summary(fun = median,geom = "line") +
        scale_x_date(labels = date_format("%Y-%m"),breaks = "1 month") +
        ggtitle("Weekly Price Evolution") +
        theme(plot.title = element_text(hjust = 0.5)) + # Adjust plot title
        theme(axis.text.x = element_text(face = "bold",color = "black",size = 8)) # Adjust x axis title

# Line chart based on daily data
ggplot(data = tab,aes(Days,Price)) +
        stat_summary(fun = median,geom = "line") +
        scale_x_date(labels = date_format("%Y-%m"), breaks = "1 month") +
        ggtitle("Daily Price Evolution") +
        theme(plot.title = element_text(hjust = 0.5)) + # Adjust plot title
        theme(axis.text.x = element_text(face = "bold",color = "black",size = 8)) # Adjust x axis title
```

Yes, price changes over time. If you compare the three time-series charts above, you will notice the following:  
1. From May to July, there was a continuous increase in retail price. This could be due to a new product launching or feature that persuaded customers to purchase within that time frame.  
2. Although retail price decreased after July for a couple months, we see another slight uptick in October. Once again, we can speculate why this occurred.  
3. The first two points can be easily observed within the monthly and weekly charts. When considering daily though, you'll notice that the highest retail price in that peak time frame of July was $540.  

**iii.** *Are prices consistent across retail outlets?*

```{r}
# Create an aggregated table showing the average Retail Price per Store.Postcode
retail_post <- aggregate(Laptops$Retail.Price, 
                    by = list(Laptops$Store.Postcode), 
                    FUN = mean, na.rm = T)
names(retail_post) <- c("Stores", "MeanRetailPrice") # Assign names

# A bar plot shows the breakout of how the stores are performing
ggplot(retail_post) + 
  geom_bar(aes(x = Stores , y = MeanRetailPrice, fill=Stores), na.rm = T, stat = "identity") +
  ggtitle("Price Consistency Per Stores") + # Title name
  ylab("Mean Retail Price") + xlab("Store")  + # Label names
  theme(plot.title = element_text(size=50)) + # Title text size
  theme_classic(base_size = 8) # Elongating the graph to show all values

# A box plot shows a little more detail about the distribution of data points, Laptops2 used from ii
boxplot(Laptops2$Retail.Price~Laptops2$Store.Postcode,data = Laptops2,
        main = "Price Consistency per Stores", xlab = "Stores", ylab ="Price",
        border = "black", las = 2, cex.axis=0.55)
```

Out of the 15 stores, only five locations are not selling at a mean of $500. Instead, they are averaging about $30 less than the top performers.

**iv.** *How does price change with configuration?*

```{r, message=FALSE, warning=FALSE}
# Add a trend line along with configuration points to show their relationship
config.smooth <- ggplot(Laptops) +
  geom_smooth(aes(x = Retail.Price, y = Configuration), color = "black") + # Labeled and colored
  geom_point(aes(x = Retail.Price, y = Configuration), color = "red") + # Labeled and colored
  labs(title = "Price Compared to Configuration", x = "Retail Price", y = "Configuration Number") + # Title and label names
  theme_classic()
config.smooth + coord_flip() # Flipped the chart to make it more readable
```

As the configuration number increases, we can see its direct effect on retail price in a positive direction. This graph shows how the two variables are closely related. 

**B** - Location Questions:  
**i.** *Where are the stores and customers located?*

```{r, message=FALSE, warning=FALSE}
# Map with open street map
map_os <- get_map(getbb("London"), source = "osm")
ggmap(map_os)

# Transform coordinates for stores locations
colnames(Laptops)
Laptops[1:5, 14:17] # overview coordinates data

# Select stores and its coordinates
stores_locations <- na.omit(unique(Laptops[, c("Store.Postcode", "store.X", "store.Y")]))
head(stores_locations)

# Create SpatialPointsDataframe
stores_locations_SP <- SpatialPointsDataFrame( 
  data = data.frame(stores_locations$Store.Postcode),    # data, stores names (postcodes)
  stores_locations[, c("store.X", "store.Y")],           # coordinates, x for "Easting", y for "Northing"
  proj4string = CRS( "+init=epsg:27700" ) )              # proj4string of the coordinates, assign CRS to data

head(stores_locations_SP)
class(stores_locations_SP)  # sp object

# Note:
## CRS, Coordinate reference systems. CRS provide a standardized way of describing locations.
## A particular CRS can be referenced by its EPSG code.
## EPSG stands for European Petroleum Survey Group. They publish a database of coordinate system 
## information plus some related documents on map projections and datums. 
## The EPSG codes are 4-5 digit numbers that represent CRS definitions. 
## proj4string is notation used in R to describe the CRS.
## EPSG:27700 Projected coordinate system for United Kingdom (UK)
## to find code p ex just google "british national grid coordinate system epsg code" 

# Transform coordinates
stores_locations_SP_LL <- spTransform(stores_locations_SP,  CRS("+init=epsg:4326")) # Data
head(stores_locations_SP_LL)

#Note:
## EPSG: 4326 code is commonly used by organizations that provide GIS data for the 
## entire globe or many countries. CRS used by Google Earth.

# Transform to data frame
stores_locations_LL <- data.frame(stores_locations_SP_LL)[,c(1:3)]      # keep first 3 columns
colnames(stores_locations_LL) <- c( "Store", "Longitude","Latitude")    # rename stores variables
head(stores_locations_LL)


#plot(Laptops$customer.X, Laptops$customer.Y,  main="Stores and Customers Locations",xlab = "Longitude",ylab = "Latitude",
#          pch = 16,cex = 0.5,col = "black",bg = "yellow") # Plot in R base with customers 'under' stores
#points(Laptops$store.X, Laptops$store.Y,pch=18, cex=2, col= "red")


store_cust <- ggplot(Laptops)+ # Overlapping customers and stores to see their locations
  geom_point(aes(customer.X, customer.Y))+
  geom_point(aes(store.X, store.Y, color = Store.Postcode,  size = 5))+ 
  theme(legend.position = "none") +
  ggtitle("Stores and Customers Locations") + # Title name
  ylab("Latitude") + xlab("Longitude")  + # Label names
  theme(plot.title = element_text(size=50)) + # Title text size
  theme_classic(base_size = 8) # Elongating the graph to show all values

store_cust

```

The majority of the customers are located within the center. If we compare this chart to the London map, we will see that most people are around the major city. The stores seem to surround that same area as well, but some are further away. This logically makes sense because in the real world, stores choose their location based on the poplation that surrounds that area.

**ii.** *Which stores are selling the most?*

```{r, message=FALSE, warning=FALSE}
L <- Laptops %>%
  ggplot(aes(x = fct_infreq(Store.Postcode))) + # Assigning frequency values to factor levels
  geom_bar() +
  labs(x = "Stores", y = "Count") + # Axis labels
  ggtitle("Stores with the Most Sales",) +  # Title name
  theme(plot.title = element_text(size=50)) + # Title size
  theme_classic(base_size = 8) # Elongating the graph to see all stores
  
L # Calls the histogram

```

There are three stores that have sales above $20,000. Those are SW1 3AU, SE1 2BN, and SW1V 4QQ.  

**iii.** *How far would customers travel to buy a laptop?*

```{r, message=FALSE, warning=FALSE}
#customercords$Store.Postcode <- Laptops$Store.Postcode
#customercords$Store.Postcode

#agg.medcustomers <- (stores_locations_LL[,c("Longitude","Latitude")], by=list(stores_locations_LL$Store.Postcode), FUN = mean)
#agg.medcustomers

#ggplot(agg.medcustomers,aes(x = Longitude,y = Latitude)) + 
#  geom_point(aes(x = Longitude, y = Latitude, color=store_postcode, fill=store_postcode), data = stores_coord_LL, shape=2, size=3) +
#  xlab("Longitude") + ylab("Latitude") + theme_classic() 

ggplot(Laptops)+ # Overlapping customers and stores to see their locations
  geom_point(aes(customer.X, customer.Y, color = Retail.Price))+
  geom_point(aes(store.X, store.Y, color = Retail.Price,  size = 5))+ 
  theme(legend.position = "none") +
  ggtitle("Stores and Customers Locations") + # Title name
  ylab("Latitude") + xlab("Longitude")  + # Label names
  theme(plot.title = element_text(size=50)) + # Title text size
  theme_classic(base_size = 8) # Elongating the graph to show all values

```

As my approach may not be correct, I believe that the logic may be slightly justified in finding how far customers would travel. When analyzing the chart above, the customers and stores are colored by retail price because if a store has a lower price point, that means that there are less customers going to that specific location. If that location has a higher retail price attached, you could assume that customers are willing to travel to that store instead.

**iv.** *Try an alternative way ok looking at how far customers traveled. Do this by creating a new column that computes the distance between customer and store.* 

```{r, message=FALSE, warning=FALSE}
Laptops2$Eucdist=sqrt((Laptops2$customer.X-Laptops2$store.X)^2+((Laptops2$customer.Y-Laptops2$store.Y)^2)) # Formula for Euclidean distance
head(Laptops2[, c(4,5,20)])  # Display the first n rows present in the input data frame
summary(Laptops2$Eucdist) # Produce result summary for Eucdist

Eucldist = densityplot(Laptops2$Eucdist, main="Distance Between the Stores and Customers (m)",xlab = "Distance (m)",col="black", )
Eucldist # Shows the distribution of distance
```

The variable 'Eucdist' was created to calculate the distance between how far customers traveled. Since there are over 142,000 observations, I wanted to focus on the main summary statistics: min, max, mean, and median. We will make an assumption that distance is measured in meters. The shortest distance is 0 meters (practically right next door), spanning to 19,892m. Additionally, when looking at the chart above, one can gather that customers were typically averaging their travel distance from 4.5 to 5 kilometers. 

**C** - Revenue Questions:  
**i.** *How do the sales volume in each store relate to Acell's revenues?*

```{r, message=FALSE, warning=FALSE}
# Bar chart shows top performers by Revenue
ggplot(data = Laptops,aes(x = Laptops$Store.Post,y = Retail.Price)) +
        geom_bar(stat = "identity") + # Type of chart
        ggtitle("Laptop Sales per Store (Revenue)") + # Title names
        xlab("Stores") + ylab("Revenue") + # Label names
        theme(plot.title = element_text(size=50)) + # Sets title size 
        theme_classic(base_size = 8) # Elongates the graph

revenue <- aggregate(Laptops$Retail.Price, # Aggregation without missing values
                    by = list(Laptops$Store.Postcode), 
                    FUN = sum, na.rm = T)
names(revenue) <- c("Stores", "Sales") # Set the name 

ggplot(revenue) + 
  geom_bar(aes(x = "", y = Sales, fill=Stores),  na.rm = T, stat = "identity", width = 1) + 
  labs(title = "Laptop Sales per Store (Revenue)", x="Stores", y="Revenue")

# Create table that shows the % and total of sales per store
revenue <- mutate(revenue,  # Adds a new variable 
                Store_wise_Sales_pct = Sales * 100 / sum(Sales)) %>%
  arrange(desc(Store_wise_Sales_pct))
revenue 

```

The top three performing stores who hold approximately 48.8% of Acell's revenue are SW1P 3AU, SE1 2BN, and SW1V 4QQ. All other stores are below $8,300,000 compared to the top three that are above $10M.

**ii.** *How does this relationship depend on the configuration?*

```{r, message=FALSE, warning=FALSE}
ggplot(Laptops) +
  geom_bar (aes(x=Store.Postcode, y=Configuration), stat="identity")  + # Set x and y
  labs(title = "Store Compared to Configuration", x="Store.Postcode", y="Configuration") + # Add graph title
  theme_classic(base_size = 8) # Elongating the graph to show all values

sales.df <- mutate(Laptops, Configuration.Small=round(Configuration/100)) # round function and call in new variable
ggplot(sales.df) +
  geom_bar(aes(x=Store.Postcode, y=Retail.Price, fill=factor(Configuration.Small)), stat="summary",
           fun='sum', position = 'fill') +
  labs(title = "Retail Price and Store Postcode", x="Store Postcode", y="Retail Price")+
  theme(axis.text.x = element_text(angle = 90))# Labeling the graph

```

You will notice that from the last exercise on Revenue versus Postcode to the bar chart above, they almost look identical. This means that the stores who have higher revenues will likely be those who are selling the higher configuration laptops. Therefore, the relationship between Configuration and Revenue based on store located is positive. 

**D** - Configuration Questions:  
**i.** *What are the details of each configuration? How does this relate to price?*  

```{r, message=FALSE, warning=FALSE}
conpric <- aggregate(Laptops$Retail.Price, by = list(Laptops$Configuration), FUN = mean, na.rm = T) #Aggregate and remove missing values

names(conpric) <- c("Configuration", "MeanRetailPrice") # Set name

ggplot(conpric) + geom_point(aes(x = Configuration, y = MeanRetailPrice), stat = "identity") # Create points with ggplot

price.config <- Laptops[,c(6:13)] # Call only those variables that are configured
price.config <- melt(price.config, "Retail.Price") # Organize the data - book reference

price.config.box <- ggplot(price.config, aes(x = value, y = Retail.Price)) + 
  geom_boxplot() + 
  stat_summary(fun = "mean", color = "red", size = .5) +
  facet_wrap(~variable, scales = "free", ncol = 3) +  # Facet wrap allows more than one variable to be shown in the same window
  labs(title = "Details of Configuration vs Price", x="Configurations", y="Retail Price") + # Label the chart
  theme_classic() 

price.config.box

```

Overall, there is a positive correlation between Configuration and MeanRetailPrice as they both increase as their values become larger. We have seen this before in the Price section. If we want to review the individual details of each configuration, you will notice the following:  
- All of the configurations increase with price the better the feature is. For example, those who bought a laptop with a HD.Size.GB 300 spent more money versus if they bought one  with 80.   
- The largest difference in price between the configurations appear to be with RAM.GB. The median price for RAM.GB 4 is about $600 while 1 and 2 are below $500.   

**ii.** *Do all stores sell all configurations?*

```{r, message=FALSE, warning=FALSE}
ggplot(Laptops, aes(y=Configuration, x=Store.Postcode)) + # Call ggplot and variables
  geom_boxplot() + # Use box plot in ggplot
  stat_summary(fun="mean",color="red", size = .5) + # A summary statistic is used to show the mean
  labs(title = "Store Configuration") + # Title name
  theme_classic(base_size = 8) # Elongating the graph to show all values # Remove background noise

number.of.stores <- count(Laptops %>%  # Creating a variable by grouping Store.Postcode 
                            group_by(Store.Postcode) %>%
                            summarize(count=n()))

number.of.config <- count(Laptops %>% # Creating a variable by grouping Configuration
                            group_by(Configuration) %>%
                            summarize(count=n()))

count.conf <- Laptops %>%  # Creating a table to show the relationship between Store and Config.
  group_by("Store Postcode" = Store.Postcode) %>%
  summarise(Number.of.same.configuration = length(unique(Configuration))) %>% 
  arrange(Number.of.same.configuration)
count.conf

```

According to the graphic above, some stores do not sell all configurations as they do not reach 864. It is a quite difficult to distinctively tell which stores those are. Therefore, in the table, you will notice that N3 1DH, E7 8NW, and KT2 5AU all fall under 750 types of configurations.
